Project Report: \\ Detecting Spam on Social Networking Sites

### 2012-05-08 docs/paper/paper.tex
paper update

@@

@@

Our ongoing observation of spam on the InterPals website provided direct intuition into the major classes of unwanted behavior that we classified as spam for this project. We noted the following broad categories:

\textbf{Advance Fee Fraud} We observed this scam to come in various flavors.  (inheritance, lottery,  etc)
\textbf{Romance Scam} 
\textbf{Sexual Solicitation}
\textbf{Porn Ads}
\textbf{Advertising External Sites}
\textbf{Webcam Ads}
\textbf{Money Muling}
\textbf{Begging \& Gift Requests}
\textbf{Business Proposals}

@@

We decided to focus on spam in private messages for this paper, given that messages account for a majority of user spam reports. In addition to using the bag-of-words representation of the message content, we aimed to identify a subset of additional, relevant ``expert" features based on public and non-public message and account information that would augment the classifier's accuracy. We first extracted a number of fields that we expected might improve classification. To choose a sample of relevant features, we computed statistics on the extracted features, both on the ham and spam corpora.

@@

\textbf{Recipient Age:} The age listed on the message recipient's profile.

\textbf{Recipient Sex:} The sex listed on the message recipient's profile.

\textbf{Recipient Country:} This was the ISO 3166-1 alpha-2 code of the recipient of the message.

@@

Due to data collection issues, recipient data (age, sex, country) was only available for the most recent 11,000 spam messages.

@@

\textbf{Sender IP:} Spammers in general had a smaller number of unique IP addresses associated with them than ham users. This is likely due to their comparatively low length of time on the site and the fact that many legitimate users are behind NATs that yield a high number of dynamic IP addresses. One possibility for future research is to investigate the ratio of unique IPs for users as a function of the time they have been registered. Likewise, examining unique /16 network blocks instead of unique IP addresses could mitigate the influence of NATs.

@@

\textbf{Sender Country:}

@@

@@

\textbf{Sender Username:}

@@

@@

Given the complexity of choosing kernels and parameters, the package includes a Python script that performs cross-validation to suggest appropriate $\gamma$ and $C$ values for the recommended radial basis function (RBF) kernel. We ran this script on a subset of 2500 ham and 2500 spam messages to obtain $C = 8$ and $\gamma = 0.0078125$.

@@

@@

@@

@@

figure \ref{fig:cprof}
figure \ref{fig:cip}
figure \ref{fig:day}
figure \ref{fig:month}
figure \ref{fig:drop}
figure \ref{fig:email}
figure \ref{fig:friends}
figure \ref{fig:usaspam}
figure \ref{fig:usahamfig:lambda}
figure \ref{fig:roclin}
figure \ref{fig:photos}
figure \ref{fig:recipage}
figure \ref{fig:sendage}
figure \ref{fig:sendsex}
figure \ref{fig:uniqip}
\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/country-prof.pdf}
    \caption{Country as stated on user profile.}
    \label{fig:cprof}
\end{figure}

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/ips-usa-spam.pdf}
    \caption{Distribution of IPs by country for spammers with profiles stating a USA location.}
    \label{fig:usaspam}
\end{figure}

@@

>>>>

\subsection{Current Anti-Spam Measures}

@@

Other anti-spam measures include widespread use of CAPTCHAs across the site and frequency caps. Short-window frequency caps are in place for all users, limiting the number of messages that can be sent per short time interval (1, 5 and 10 minutes). New users are also subject to a per-day cap on the number of unique users with whom they are able to initiate contact.

Users are able to filter messages by age, sex, continent and country. At this time,

@@

In the course of collecting the data, we observed a number of different types of undesirable messages. Our ongoing observation of spam on the InterPals website provided direct intuition into the major classes of unwanted behavior that we classified as spam for this project. We noted the following broad categories:

\textbf{Advance Fee Fraud} We observed this scam to come in various flavors.  (inheritance, lottery,  etc)
\textbf{Romance Scams:} These 
\textbf{Sexual Solicitation}
\textbf{Porn Ads}
\textbf{Advertising External Sites}
\textbf{Webcam Ads}
\textbf{Money Muling}
\textbf{Begging \& Gift Requests}
\textbf{Business Proposals}

@@

We decided to focus on spam in private messages for this paper, given that messages account for a majority of user spam reports. In addition to using the bag-of-words representation of the message content, we aimed to identify a subset of additional, relevant ``expert" features based on public and non-public message and account information that would augment the classifier's accuracy. We first extracted a number of fields that we expected might improve classification.

To choose a sample of relevant features, we computed statistics on the extracted features, both on the ham and spam corpora. These statistics were generated using SQL on a table created from the merged and cleaned data generated in the process described above. SQL makes it trivial to calculate means and standard errors, but medians are more resource-intensive, so we omitted them from our

@@

\textbf{Recipient Age/Sex/Country:} These fields were the age, sex and ISO 3166-1 alpha-2 code of the message 
recipient, as listed on their profile. Unfortunately, due to data collection issues, this data was only available 
for the most recent 11,000 spam messages. While the male-to-female ratio of the recipients was very close to that 
of non-spam recipients (both were 47\% male to 53\% female), we found that the age of recipients was typically higher.

@@

@@

\textbf{Sender IPs:} Spammers in general had a smaller number of unique IP addresses associated with them than ham users, as per figure \ref{fig:uniqip}. The mean for spammers was 11.82 with a standard deviation of 52.52, whereas ham users saw a mean of 175.51 and standard deviation of 322.19. This is likely due to their comparatively low length of time on the site and the fact that many legitimate users are behind NATs that yield a high number of dynamic IP addresses. One possibility for future research is to investigate the ratio of unique IPs for users as a function of the time they have been registered. Likewise, examining unique /16 network blocks instead of unique IP addresses could mitigate the influence of NATs.

@@

\textbf{Sender Age:} Spam user age

\textbf{Sender Country:} The distribution of countries as stated by users on their profile and as revealed by their IP address differed remarkably between the spam and ham user groups. 48\% of spam users claimed to be from the United States, with the next most popular countries being UK, Ghana, Senegal, Germany and Canada taking combined 25\%. In contrast, only 15\% ham users claimed to be from the United States, with Russia, Korea, UK, France and Germany combining for 24\%.

figure \ref{fig:usaspam}
figure \ref{fig:usahamfig:lambda}
\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/ips-usa-spam.pdf}
    \caption{Distribution of IPs by country for spammers with profiles stating a USA location.}
    \label{fig:usaspam}
\end{figure}

@@

We also noticed that Afghanistan and Albania featured prominently at the 11th and 15th most popular countries f......

@@

\textbf{Username, Name:} While we collected these fields, we did not end up processing or analyzing them further. That said, we saw considerable repetition in the usernames of scammers and it seems that analyzing the substrings in these fields could yield useful features.

@@

Support Vector Machines are a classification method  \cite{boser}, \cite{cortesv95}

@@

Given the complexity of choosing kernels and parameters, the package includes a Python script that performs cross-validation to suggest appropriate $\gamma$ and $C$ values for the recommended radial basis function (RBF) kernel. The RBF kernel is a real-valued function such that

\begin{itemize}
\item $K(\mathbf{x}_i, \mathbf{x}_j) = (\gamma\mathbf{x}_i^\intercal\mathbf{x}_j+r)^d, \gamma>0$
\end{itemize}
We ran this script on a subset of 2500 ham and 2500 spam messages to obtain $C = 8$ and $\gamma = 0.0078125$.

@@

\subsection{Spam Grouping}

As a part of this project, we performed some qualitative analysis on spam and differentiated between the major categories of undesirable messages found on this SNS. However, we

\subsection{Features}

@@

It is also anecdotally clear from moderator log messages that spammers re-use their profile photos and these photos are often of celebrities, porn stars or simply stock photos. Hashing these images

@@

\subsection{Testing \& Evaluation}

- test SVM and other classifiers on newer data
-

@@

figure \ref{fig:cprof}
figure \ref{fig:cip}
figure \ref{fig:day}
figure \ref{fig:month}
figure \ref{fig:drop}
figure \ref{fig:email}
figure \ref{fig:friends}
figure \ref{fig:roclin}
figure \ref{fig:photos}
figure \ref{fig:recipage}
figure \ref{fig:sendage}
figure \ref{fig:sendsex}
figure \ref{fig:roc-25}
\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/country-prof.pdf}
    \caption{Country as stated on user profile.}
    \label{fig:cprof}
\end{figure}

@@

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/roc-25.pdf}
    \caption{LIBSVM ROC curves for classifier using bag-of-words and bag-of-words with ``expert" features on 50k messages.}
    \label{fig:roc-25}
\end{figure}
